FraCaS: Temporal Analysis

### 2020-06-09 paper2020/paper.tex
handle various todos

\paragraph{Temporal Semantics}
\jp{More general intro to temporal semantics? Should this be moved
  somewhere else?}

We use notions of ``stative'', ``achievement'', ``activity''. Our
definition do not seem to match perfectly the definitions found in the
literature; In general, we use the term achievement to include
accomplishment as well. Insofar as the test suite is concerned, we
find that these two categories can be collapsed into a single one.  As
such, we think it is a good classification, given that they generally
afford the correct inferences for the testsuite. (We assume the
testsuite is correct.)

@@

\onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
\\
{\small\begin{multline*}
\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
\end{multline*}
}}
In our semantics we deal only with present and simple past
tenses. \jp{Is this enough for theory?} Indeed we find that FraCas
does not exercise additional specific tenses (if a more complicated
tense is used, the additional information is also carried by adverbs
or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.

The procedure to handle tense annotations is as follows:
\begin{itemize}
\item If the tense is the past, and the temporal context is
  unspecified, then we locally quantify over a time interval
  $[t_0,t_1]$, such that $t_1 < now$, where $now$ is a logical
  constant representing the current timepoint. The temporal context
  then becomes this interval.
\item If the tense is the present and the temporal context is
  unspecified, then the temporal context becomes the simple
  $(now,now)$ interval.
\item If the temporal context is specified (for example due to the presence of an adverb or an
  averbial clause, such as ``before James swam''), then the tense does
  not create a new interval, but it may constrain it. Typically, a
  past tense adds the constraint that the temporal context ends before
  the timepoint $now$.
\end{itemize}

@@

Unicity of action holds only if the aspectual class of the proposition
$P$ is \emph{achievement} or \emph{actitivty}.\todo{reference in literature}

(The difference between actitivty and achievement is that for the
latter time intervals can be assumed to be of nil duration. In reality, this is an oversimplification as achievements are usually of short duration. but not nil. However, 
this appears to play little role, if any, in our analysis, and given the task-driven nature of it as well.)

@@

\[\begin{array}{l}
∃x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃y. novel(y) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Our analysis does not need to treat this last case specially. Indeed,
even if $write(smith,x,.,.)$ is an activity \todo{This should be an accomplishment and thus, achievement in our classification} and thus subject to
unicity of action, in (280), $x$ is quantified existentially; we have
two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot deduce equality of the
intervals $t_1,t_2$ and $t_3,t_4$. In turn, the hypothesis cannot be refuted.

@@

Finally we note that not all verbs are subject to the
stative/achievement distinction induced by the progressive. For
example, the phrases ``John ran'' and ``John is running'' appear to be
logically equivalent, for our purposes\todo{For the FraCaS suite?}.

>>>>

\paragraph{Temporal Semantics}

In this paper we borrow several notions from classical temporal
semantics such as ``stative'', ``achievement'', ``activity'', etc.,
even though our definitions do not perfectly match the classical
ones. We explain our precise meaning forthese terms in the body of the
paper. Nevertheless, we refer the reader to \citet{steedman:2000} for
an extensive review of temporal semantics.

For the \textit{cognoscenti}, we can already point out some
differences in terminology: we use the term activity as a general term
which encompasses the three classical notions of activites,
achievements and accomplishments. Indeed, insofar as the test suite is
concerned, we find that these three categories can be collapsed into a
single one.  As such, we think it is a good classification, given that
they generally afford the correct inferences for the testsuite. (In
this paper, we always assume that the FraCaS testsuite is correct.)

@@

\onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
\\
{\small\begin{multline*}
\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
\end{multline*}
}}
In our semantics we deal only with present and past
tenses (simple and continuous). Indeed we find that FraCas
does not exercise additional specific tenses (if a more complicated
tense is used, the additional information is also carried by adverbs
or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.

Even though we will discuss a refinement to handle the past continuous
at the end of this section, the procedure to handle tense annotations
is as follows:
\begin{itemize}
\item If the tense is the past, and the temporal context is
  unspecified, then we locally quantify over a time interval
  $[t_0,t_1]$, such that $t_1 < now$, where $now$ is a logical
  constant representing the current timepoint. The temporal context
  then becomes this interval.
\item If the tense is the present and the temporal context is
  unspecified, then the temporal context becomes the simple
  $(now,now)$ interval.
\item If the temporal context is specified (for example due to the presence of an adverb or an
  averbial clause, such as ``before James swam''), then the tense does
  not create a new interval, but it may constrain it. Typically, a
  past tense adds the constraint that the temporal context ends before
  the timepoint $now$.
\end{itemize}

@@

Unicity of action holds only if the aspectual class of the proposition
$P$ is \emph{actitivty}~\citep{steedman:2000} (which, for our
purposes, includes \emph{achievements} and \emph{accomplishments}).

(The difference between actitivty and accomplishments on the one and
achievement on the other hand is that for the latter, time intervals
can be assumed to be of nil duration. In reality, this is an
oversimplification as achievements are usually of short duration, but
not nil. However, this appears to play little role, if any, in our
analysis, and given the its task-driven nature as well.)

@@

\[\begin{array}{l}
∃x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃y. novel(y) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,y,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Our analysis does not need to treat this last case specially. Indeed,
even if $write(smith,x,.,.)$ is an activity and thus subject to
unicity of action, in (280), $x$ is quantified existentially; we have
two \emph{different} actions: $write(smith,x,t_1,t_2)$ and
$write(smith,y,t_3,t_4)$, because $x \neq y$, and thus we cannot
deduce equality of the intervals $t_1,t_2$ and $t_3,t_4$. In turn, the
hypothesis cannot be refuted.

@@

Finally we note that not all verbs are subject to the
stative/achievement distinction induced by the progressive. For
example, the phrases ``John ran'' and ``John is running'' appear to be
logically equivalent, for entailment purposes.
